---
# ==============================================================================
# QUARTO PRESENTATION TEMPLATE
# Four Uses of Regression — Teaching Presentation
#
# Fill in the placeholders marked [LIKE THIS].
# Read the comments (<!-- ... -->) before each slide. They explain the
# purpose of the slide, not just the content. Delete the comments once
# you have understood them — they should not appear in your final file.
#
# The structure is fixed. Do not add or remove slides unless absolutely necessary.
# ==============================================================================

title: "[Your Title: an Assertion, Not a Label]"
subtitle: "[Topic N: e.g., 'Regression for Description']"
author: "[Your Name] and [Your Partner]"
date: today

format:
  beamer: default

execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r setup}
# Load only what you need. No kitchen-sink package imports.
library(tidyverse)
library(broom)     # tidy() for regression output
library(ggplot2)

# Set a seed so your simulation is reproducible. Use your birthday,
# a favorite number, whatever — but set one.
set.seed(20260421)

# A minimalist ggplot theme. Remove gridlines you are not using.
theme_set(theme_minimal(base_size = 16))
```

<!--
===============================================================================
TITLE CONVENTION REMINDER

Every slide title below is currently a LABEL (e.g., "The Question").
That is placeholder text. Before you submit, rewrite every title as an
ASSERTION — a one-sentence claim about what the slide demonstrates.

BAD:  "The Broken Case"
GOOD: "Adjusting for a collider inverts the sign of the gap"

BAD:  "Diagnosis"
GOOD: "The failure is wrong comparisons, not linearity violations"

If someone reads only your slide titles in sequence, they should
understand your whole argument. Test this before you submit.
===============================================================================
-->

## The Question <!-- Change this title to be an assertion -->

<!--
SLIDE 1 PURPOSE: Establish what question your presentation answers.
ONE sentence. No jargon. The audience should understand the stakes
before they see any math.

Examples of good Slide 1 claims, by topic:
  Topic 1: "Does running a regression clarify or hide the comparison
           you are making?"
  Topic 2: "What makes a regression a good description, and how would
           you know?"
  Topic 3: "When does a sample regression coefficient tell us something trustworthy
           about the population?"
  Topic 4: "Under what assumptions does a regression coefficient tell us something trustworth about a
           causal effect?"

Do NOT use this slide for "Today we will discuss..." That is not a
question, it is an agenda.
-->

[Your one-sentence question or claim goes here.]

---

## What Success Looks Like

<!--
SLIDE 2 PURPOSE: Define what it means for regression to SUCCEED at
your goal. Do this BEFORE you show any method. You need a criterion
against which to judge success and failure.

This is the hardest slide. If you cannot define success precisely,
you cannot show that anything works or fails. Spend time here.

Keep it to 2-4 bullets or a single clear statement. Do not write a
paragraph. The audience should leave this slide knowing the standard
to which you will be holding regression given your particular purpose.
-->

- [Criterion 1]
- [Criterion 2]
- [Criterion 3, if needed]

---

## The Working Case: Setup

<!--
SLIDE 3 PURPOSE: Show how you coded up the simulation for the case where
regression succeeds. The audience needs to see:
  1. What variables exist
  2. How they are related (ideally as a diagram or equations)
  3. Why this data-generating process/simulation represents the case
  where regression succeeds at the use you are evaluation. one where your goal
  is achievable

DO NOT paste the full simulation code. Show the structure, not the syntax. If
you want to show R code, show the ESSENTIAL lines with everything else
suppressed. You can also show pseudo-code.

-->

```{r working-dgp, eval=TRUE}
#| echo: true
#| code-line-numbers: false

# SIMULATE THE WORKING CASE
# Edit this to generate data where regression DOES what you want.
# [Your simulation code here. Keep it simple enough to fit on the slide.]
# Example structure:
```

[Briefly describe what this data-generating process represents — e.g., "Random
binary treatment, one covariate observed, binary outcome; observational study
with treatment caused by two observed covariates and an unobserved confounder,
continuous outcome that also depends on those covariates plus an unobserved
confounder."]

---

## The Working Case: Regression Does the Job

<!--
SLIDE 4 PURPOSE: Show the regression output and demonstrate that it
achieved your goal (the criterion from Slide 2).

DO NOT dump a full regression table with 15 coefficients. Show:
  - The ONE number that matters (the coefficient, SE, p-value, or two numbers of the CI, false positive rate (Type I error rate), coverage rate,
    , mean-squared error, bias), and
  - A visual if possible (estimate vs. truth, CI covering the true
    value, distribution of estimates across simulations)

If the audience cannot see the success without you narrating, the
slide is too dense.
-->

```{r working-result}
#| fig-cap: "[Describe what the plot shows in one line]"
```

**Key finding:** [One sentence. E.g., "The coefficient bounces around the true value and is not systematically too high or too low ; CI coverage is 94%; the comparison by hand that we desire is exactly the comparison the regression provides."]

---

## The Broken Case: Setup

<!--
SLIDE 5 PURPOSE: Show the DGP for the case where regression FAILS.

The difference between this DGP and the working-case DGP should
usually be ONE thing. You are breaking one assumption cleanly, not
six at once. A muddy failure teaches nothing.

Tell the audience explicitly what you are changing:
  - "Same as before, but now the confounder is unmeasured."
  - "Same DGP, but we now select into the sample based on Y."
  - "Same structure, but X is a collider of Z and Y, not a confounder."
-->

```{r broken-dgp, eval=TRUE}
#| echo: true

# SIMULATE THE BROKEN CASE
# Change ONE thing from the working DGP. State what you changed.

# [Your broken DGP here.]
```

**What changed:** [One sentence describing the single modification.]

---

## The Broken Case: Regression Fails

<!--
SLIDE 6 PURPOSE: Show that the regression fails, in the SAME way
you showed it succeeded on Slide 4. Parallel structure matters —
the audience should be able to compare Slide 4 and Slide 6 directly.

The failure must be VISIBLE. Do not say "the estimate is biased"
and show a number. Show the estimate next to the truth, or show
coverage collapsing, or show the sampling distribution drifting
away from the true value.
-->

```{r broken-result}
#| fig-cap: "[Describe what the plot shows]"

# Fit the SAME model to the broken data.
# Show the same diagnostic you showed in Slide 4 — now failing.
```

**Key finding:** [One sentence. E.g., "The coefficient is biased by 10 point; CI coverage drops to 12%."]

---

## Why It Broke

<!--
SLIDE 7 PURPOSE: Diagnose the failure. This is the conceptual
heart of your presentation. Without a clear diagnosis, the broken
case is just a demonstration — not a lesson.

Name the assumption that was violated. Use the vocabulary of your
topic, for example:
  Topic 1 or 2: specification searching, forking paths, p-hacking
  Topic 1 or 2: confounder vs. mediator vs. collider; over-adjustment
  Topic 3: non-random sampling, clustered SE, selection into sample, samples too small for CLT, outcome too skewed or rare binary for CLT
  Topic 4: unmeasured confounding, collider bias, post-treatment bias, forbidden comparisons

A good diagnosis is at the level of the ASSUMPTION, not the symptom. "The
overall effect was negative even though when we looked at the lmw or TWFE
weights by hand, all of the strata have positive effects." is a symptom not a diagnosis.

-->

The assumption violated: **[name it precisely]**

[One or two sentences explaining, in the topic's native vocabulary, why this violation produced the failure you saw on Slide 6.]

---

## The Fix (or the Honest Admission)

<!--
SLIDE 8 PURPOSE: Can the failure be fixed with more data, a better
specification, or a different design? Or is it unfixable given what
we have in the simulated data?

BE HONEST. Some failures have no remedy short of a different study.
If yours is one of those, say so. Intellectual honesty on this slide
builds ethos more than a false solution does.

Do not invent a fix for the sake of symmetry. "There is no way to
recover the causal effect from this data" is a legitimate and
correct answer for many Topic 4 failures.
-->

[One of:]

- **Fixable by** [cluster-robust SEs / DAG-based adjustment / IV / etc.]
- **Unfixable with this data** because [reason]

---

## The Boundary

<!--
SLIDE 9 PURPOSE: State the conditions under which regression DOES
and DOES NOT serve your goal. This generalizes beyond your specific
simulation.

The audience should leave this slide knowing when they should trust
a regression for your topic and when they should be skeptical.
This is the transferable lesson.
-->

Regression serves [your goal] when:

- [Condition 1]
- [Condition 2]

It fails when:

- [Failure condition 1]
- [Failure condition 2]

---

## Takeaway

<!--
SLIDE 10 PURPOSE: Give the audience ONE sentence they will remember
tomorrow. Not a summary. Not a thank-you. A takeaway.

Return to the question from Slide 1 — now answered. If you cannot
close the loop between Slide 1 and Slide 10, your argument has
drifted. Go back and fix it.

DO NOT end with "Questions?" That is lazy. Put your takeaway up,
let it sit while you take questions, and let people leave with
your sentence in their heads.
-->

# [Your one-sentence takeaway.]

<!-- ============================================================ -->
<!-- PRE-SUBMISSION CHECKLIST                                     -->
<!--                                                              -->
<!-- Before you submit, walk through your deck and confirm:       -->
<!--                                                              -->
<!-- [ ] Every slide title is an ASSERTION, not a label           -->
<!-- [ ] Every slide expresses exactly ONE idea                   -->
<!-- [ ] No slide is a wall of sentences                          -->
<!-- [ ] The working case and broken case differ in ONE thing     -->
<!-- [ ] The failure on Slide 6 is visible, not just asserted     -->
<!-- [ ] Slide 7 names the assumption that was violated           -->
<!-- [ ] Slide 10 closes the loop with Slide 1                    -->
<!-- [ ] You have rehearsed and it fits in 10 minutes             -->
<!-- [ ] You have deleted every instructional comment in this     -->
<!--     template before rendering the final version              -->
<!-- ============================================================ -->
